---
title: "Cyclistic bike-share analysis case study"
author: "Fernanda Alfian"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Scenario

You are a junior data analyst working in the marketing analyst team at Cyclistic, a bike-share company in Chicago. The directorof marketing believes the company’s future success depends on maximizing the number of annual memberships. Therefore, your team wants to understand how casual riders and annual members use Cyclistic bikes diﬀerently. From these insights, your team will design a new marketing strategy to convert casual riders into annual members. But ﬁrst, Cyclistic executives must approve your recommendations, so they must be backed up with compelling data insights and professional data visualizations. 



## Setting up my packages

We need to set the dictionary to helps us to input, process, and visualize data.
We need to install some packages like tidyverse, ggplot2, dplyr, readr.


This package will help to wragle the data
```{r tidyverse package, echo=FALSE}
#install.packages("tidyverse")
library("tidyverse")
```

This package will help to visualize the data
```{r ggplot2 package, echo=FALSE}
#install.packages("ggplot2")
library("ggplot2")
```

This package will help to wragle the data
```{r dplyr package, echo=FALSE}
#install.packages("dplyr")
library("dplyr")
```

This package will help to read the .csv files data
```{r readr package, echo=FALSE}
#install.packages("readr") 
library("readr")
```

```{r}
library("lubridate")
```


## Collect the data
#### Import the .csv files data

We import 13 .csv files data, from January 2021 until Januari 2022 from local computer. 
Rename for each data into format : Month_Year 

The data are downloaded from [this link](https://divvy-tripdata.s3.amazonaws.com/index.html)

```{r import .csv files data}
Jan_2021 <- read.csv("C:\\Users\\ferna\\OneDrive\\Documents\\Case Study Coursera\\divvy-tripdata\\202101-divvy-tripdata.csv")
Feb_2021 <- read.csv("C:\\Users\\ferna\\OneDrive\\Documents\\Case Study Coursera\\divvy-tripdata\\202102-divvy-tripdata.csv")
Mar_2021 <- read.csv("C:\\Users\\ferna\\OneDrive\\Documents\\Case Study Coursera\\divvy-tripdata\\202103-divvy-tripdata.csv")
Apr_2021 <- read.csv("C:\\Users\\ferna\\OneDrive\\Documents\\Case Study Coursera\\divvy-tripdata\\202104-divvy-tripdata.csv")
May_2021 <- read.csv("C:\\Users\\ferna\\OneDrive\\Documents\\Case Study Coursera\\divvy-tripdata\\202105-divvy-tripdata.csv")
Jun_2021 <- read.csv("C:\\Users\\ferna\\OneDrive\\Documents\\Case Study Coursera\\divvy-tripdata\\202106-divvy-tripdata.csv")
Jul_2021 <- read.csv("C:\\Users\\ferna\\OneDrive\\Documents\\Case Study Coursera\\divvy-tripdata\\202107-divvy-tripdata.csv")
Aug_2021 <- read.csv("C:\\Users\\ferna\\OneDrive\\Documents\\Case Study Coursera\\divvy-tripdata\\202108-divvy-tripdata.csv")
Sep_2021 <- read.csv("C:\\Users\\ferna\\OneDrive\\Documents\\Case Study Coursera\\divvy-tripdata\\202109-divvy-tripdata.csv")
Oct_2021 <- read.csv("C:\\Users\\ferna\\OneDrive\\Documents\\Case Study Coursera\\divvy-tripdata\\202110-divvy-tripdata.csv")
Nov_2021 <- read.csv("C:\\Users\\ferna\\OneDrive\\Documents\\Case Study Coursera\\divvy-tripdata\\202111-divvy-tripdata.csv")
Dec_2021 <- read.csv("C:\\Users\\ferna\\OneDrive\\Documents\\Case Study Coursera\\divvy-tripdata\\202112-divvy-tripdata.csv")
Jan_2022 <- read.csv("C:\\Users\\ferna\\OneDrive\\Documents\\Case Study Coursera\\divvy-tripdata\\202201-divvy-tripdata.csv")
```


## Data wrangling and combining into single file

We need to look that all files has the same column for each file. 
If there are some files has different column, we need to make the column name has same standard like other data. 

```{r column Jan_2021}
colnames(Jan_2021)
```

```{r column Feb_2021}
colnames(Feb_2021)
```

```{r column Mar_2021}
colnames(Mar_2021)
```

```{r column Apr_2021}
colnames(Apr_2021)
```

```{r column May_2021}
colnames(May_2021)
```

```{r column Jun_2021}
colnames(Jun_2021)
```

```{r column Jul_2021}
colnames(Jul_2021)
```

```{r column Aug_2021}
colnames(Aug_2021)
```

```{r column Sep_2021}
colnames(Sep_2021)
```

```{r column Oct_2021}
colnames(Oct_2021)
```

```{r column Nov_2021}
colnames(Nov_2021)
```

```{r column Dec_2021}
colnames(Dec_2021)
```

```{r column Jan_2022}
colnames(Jan_2022)
```

Note: All of files has same column. Now, we can move to next step.




We want to looking for the data type for each files. We need to make sure that for each files has same data type. 
That helps us to change the data type together if it needed

```{r str Jan_2021}
str(Jan_2021)
```

```{r str Feb_2021}
str(Feb_2021)
```

```{r str Mar_2021}
str(Mar_2021)
```

```{r str Apr_2021}
str(Apr_2021)
```

```{r str May_2021}
str(May_2021)
```

```{r str Jun_2021}
str(Jun_2021)
```

```{r str Jul_2021}
str(Jul_2021)
```

```{r str Aug_2021}
str(Aug_2021)
```

```{r str Sep_2021}
str(Sep_2021)
```

```{r str Oct_2021}
str(Oct_2021)
```

```{r str Nov_2021}
str(Nov_2021)
```

```{r str Dec_2021}
str(Dec_2021)
```

```{r str Jan_2022}
str(Jan_2022)
```


From the results, we can see that started_at and ended_at column has char data type. We need to convert the data type char to date. 
Because all the columns name are on the same position, we can change the data type together in single file not individualy, after combining all of files into single file.

```{r combining to single file}
all_trips_data <- rbind(Jan_2021, Feb_2021, Mar_2021, Apr_2021, May_2021, Jun_2021, Jul_2021, Aug_2021, Sep_2021, Oct_2021, Nov_2021, Dec_2021, Jan_2022)

```

look the column name
```{r}
head(all_trips_data)
```

look for the data type
```{r}
str(all_trips_data)
```


Change the data type char to date on started_at and ended_at column
```{r change data type }
all_trips_data <- mutate(all_trips_data,
                         started_at = as.POSIXct(started_at),
                         ended_at = as.POSIXct(ended_at)
                         )
```


Look if the data type has change
```{r}
str(all_trips_data)
```


Looking for summary of all_trips_data
```{r summary all_trips_data}
summary(all_trips_data)
```


We can see the summary of the all_trips_data.
But these are not that we're looking for.

There are several problems for this data. 
(1) We want to see the date, month, hour of trip, day, day of week. We create several columns.
(2) We want to see the trip length of each trip. We create a new column to calculate the different between ended_et and started_at.
(3) We want to remove some data that shows negative value of trip length.
(4) Some of stations are warehouse and testing of divvy bikes and we need to clean it up from the data.
(5) We want to see if the rides return the shared bike on the same station as the started station We create a column to see it.

All of these problems, we can see the solutions from next section.




## Data cleaning and data exploration

(1) We want to see the date, month, hour of trip, day, day of week. We create several columns.
```{r}
all_trips_data$date <- as.Date(all_trips_data$started_at)
all_trips_data$month <- format(as.Date(all_trips_data$started_at), "%m")
all_trips_data$trip_hour <- (format(all_trips_data$started_at, "%H"))
all_trips_data$day <- format(as.Date(all_trips_data$started_at), "%w")
all_trips_data$day_of_Week <- format(as.Date(all_trips_data$started_at), "%A")
all_trips_data$datetime <- (format(all_trips_data$started_at, "%H:%M:%S"))
```


(2) We want to see the trip length of each trip. We create a new column to calculate the different between ended_et and started_at.
```{r}
all_trips_data$ride_length <- difftime(all_trips_data$ended_at, all_trips_data$started_at)
```


We checked the data column and data type is correct, and we can move to the next step.
```{r}
head(all_trips_data)
str(all_trips_data)
```


The ride_length data type is char, is not that we want to see. 
We need to convert the data type to numeric, because if the data type is char, we can't calculate the column for next step analysis. 

```{r}
is.numeric(all_trips_data$ride_length)
```


```{r}
all_trips_data$ride_length <- as.numeric(as.character(all_trips_data$ride_length))
is.numeric(all_trips_data$ride_length)
```
The data type is changed to numeric value




(3) We want to remove some data that shows negative value of trip length.
(4) Some of stations are warehouse and testing of divvy bikes and we need to clean it up from the data.


From this step, we create a new data frame, just want to see the number of observation. That can help us to see the percentage of number of observation will affect the main data frame if the number of obs in new data frame are deleted in main data frame.


```{r ride_length negative}
all_trips_data_ride_length_negative <- all_trips_data[(all_trips_data$ride_length<0),]
```
From this data frame, we can see total obs, and from this code it will be our reference for ride_length negative in main data_frame


```{r latitude and longitute are N/A}
all_trips_na_lat_long <- all_trips_data[(all_trips_data$start_lat == "" | is.na(all_trips_data$start_lat) | all_trips_data$start_lng == "" | is.na(all_trips_data$start_lng) | all_trips_data$end_lat == "" | is.na(all_trips_data$end_lat) |  all_trips_data$end_lng == "" | is.na(all_trips_data$end_lng)),]
```
From this code / data frame, tell us about the latitude and longitute that has N/A values in data, and we can delete the data from main data frame.



```{r N/A stations}
all_trips_na_stations <- all_trips_data [(all_trips_data$start_station_name == ""| is.na(all_trips_data$start_station_name) | all_trips_data$end_station_name == "" | is.na(all_trips_data$end_station_name)),]
```
There are a lot of data from N/A stations. 
From [divvy information website](https://divvybikes.com/pricing/day-pass), for day pass they have more flexibility to return the bike. We assume that they can just lock the bike in anywhere, not only on station.

From the data, there are some station that indicate the warehouse / testing bike of divvy station, and we delete it later.





We create a new data frame called all_trips_data_v2, to store our clean data.

```{r}
all_trips_data_v2 <- all_trips_data[!(all_trips_data$ride_length<0),]
```

```{r}
all_trips_data_v2 <- all_trips_data_v2[!(all_trips_data_v2$start_lat == "" | is.na(all_trips_data_v2$start_lat) | all_trips_data_v2$start_lng == "" | is.na(all_trips_data_v2$start_lng) | all_trips_data_v2$end_lat == "" | is.na(all_trips_data_v2$end_lat) |  all_trips_data_v2$end_lng == "" | is.na(all_trips_data_v2$end_lng)),]
```


```{r}
all_trips_data_v2 <- all_trips_data_v2 [!((all_trips_data_v2$start_station_name == "WATSON TESTING - DIVVY" & !is.na(all_trips_data_v2$start_station_name)) | (all_trips_data_v2$start_station_name == "HUBBARD BIKE-CHECKING (LBS-WH-TEST)" & !is.na(all_trips_data_v2$start_station_name)) | (all_trips_data_v2$start_station_name == "Base - 2132 W Hubbard Warehouse" & !is.na(all_trips_data_v2$start_station_name)) | (all_trips_data_v2$start_station_name == "Hubbard Bike-checking (LBS-WH-TEST)" & !is.na(all_trips_data_v2$start_station_name)) | (all_trips_data_v2$start_station_id == "Hubbard Bike-checking (LBS-WH-TEST)" & !is.na(all_trips_data_v2$start_station_id))),]
```

`


(5) We want to see if the rides return the shared bike on the same station as the started station We create a column to see it.
```{r}
all_trips_data_v2 <- all_trips_data_v2 %>% 
  mutate(start_end_same_station = case_when(
    all_trips_data_v2$start_station_name == all_trips_data_v2$end_station_name ~ "Yes",
    all_trips_data_v2$start_station_name != all_trips_data_v2$end_station_name ~ "No"))

```

We want to see the data type from new data frame
```{r}
str(all_trips_data_v2)
```


We want to change the trip_length from second format to minute rounded
```{r}
all_trips_data_v2$ride_length_min <- all_trips_data_v2$ride_length/60
all_trips_data_v2$ride_length_min <- round(all_trips_data_v2$ride_length_min, digits=0)
```


## Desctiptive Analysis



We want to see the number of rides for member and casual 

```{r number of rides, echo=FALSE}
all_trips_data_v2 %>%
  group_by(member_casual) %>%
  dplyr::summarise(count=n())
```


We want to see the mean, meadian, max of ride_length
```{r summary data, echo=FALSE}
summary(all_trips_data_v2$ride_length_min)
```


Compare ride_length for member and casual
```{r echo=FALSE}
aggregate(all_trips_data_v2$ride_length_min ~ all_trips_data_v2$member_casual, FUN = mean) #mean
aggregate(all_trips_data_v2$ride_length_min ~ all_trips_data_v2$member_casual, FUN = median) #median
aggregate(all_trips_data_v2$ride_length_min ~ all_trips_data_v2$member_casual, FUN = max) #max
aggregate(all_trips_data_v2$ride_length_min ~ all_trips_data_v2$member_casual, FUN = min) #min
```




We want to see the ride_length per day and for member and casual
```{r echo=FALSE}
aggregate(all_trips_data_v2$ride_length_min ~ all_trips_data_v2$member_casual + all_trips_data_v2$day_of_Week, FUN = mean)
```

Because the day_of_week not shorted acending. We need to short the day_of_week acending order using levels 
```{r include=FALSE}
all_trips_data_v2$day_of_Week <- ordered(all_trips_data_v2$day_of_Week, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

```

We can see the day_of_week is shorted acending
```{r echo=FALSE}
aggregate(all_trips_data_v2$ride_length_min ~ all_trips_data_v2$member_casual + all_trips_data_v2$day_of_Week, FUN = mean)
```




We want to see the total of rides minutes for member and casual
```{r echo=FALSE}
aggregate(all_trips_data_v2$ride_length_min ~ all_trips_data_v2$member_casual, FUN = length)
```



We want to see the ride length for ridable type for member and casual
```{r echo=FALSE}
aggregate(all_trips_data_v2$ride_length_min ~ all_trips_data_v2$rideable_type + all_trips_data_v2$member_casual, FUN = length)
```



We want to see the ride length for each days and member and casual
```{r echo=FALSE}
aggregate(all_trips_data_v2$ride_length_min ~ all_trips_data_v2$member_casual + all_trips_data_v2$day_of_Week, FUN = length)
```



We want to see the rides who start and ended the rides in same or other station
```{r echo=FALSE}
aggregate(all_trips_data_v2$ride_length_min ~ all_trips_data_v2$member_casual + all_trips_data_v2$start_end_same_station, FUN = length)
```











## Data Visualization

Because the visual in Y axis has exponential number, we want to change to normal number
```{r include=FALSE}
options(scipen = 999) 
```


### Rides by member type (member and casual)
#### Number of rides by member type

Looking for the number of rides by member type
```{r echo=FALSE}
all_trips_data_v2 %>% 
  group_by(member_casual) %>% 
  dplyr::summarize(number_of_rides = n()
                   ,average_duration = mean(ride_length_min)) %>% 
  arrange(member_casual) %>% 
  ggplot(aes(x = member_casual, y = number_of_rides, fill = member_casual)) +
  geom_col(position = "dodge")
```

Count the exact of rides by member type
```{r echo=FALSE}
all_trips_data_v2 %>%
  group_by(member_casual) %>%
  dplyr::summarise(count=n())
```

As we can see above, the number of annual member are higher than the casual riders. There are total 3149602	or 55.31% riders of annual member while for the casual member, 2543902 or 44.68% riders.


#### average_durations of rides by member type
Looking for the average duration in minute of rides by member type
```{r echo=FALSE}
all_trips_data_v2 %>% 
  group_by(member_casual) %>% 
  dplyr::summarize(number_of_rides = n()
                   ,average_duration = mean(ride_length_min)) %>% 
  arrange(member_casual) %>% 
  ggplot(aes(x = member_casual, y = average_duration, fill = member_casual)) +
  geom_col(position = "dodge")
```

Count the exact average_durations of rides by member type
```{r echo=FALSE}
all_trips_data_v2 %>% 
  group_by(member_casual) %>% 
  dplyr::summarize(ride_length_min = mean(ride_length_min))
```



Otherwise from the number of rides. 
Average duration for Casual riders are higher than annual member. 
From the data above, we can assume that casual riders rent the bike to cycle around for quite a long distance, but for annual members only bikes to commute to work each day.

But we need more data to accept that assumtion









### Rides by rideable_type
#### Number of rides by rideable_type and member type

Run the data visualization of number_of_rides by rideable_type and member type
```{r echo=FALSE}
all_trips_data_v2 %>% 
  group_by(member_casual, rideable_type) %>% 
  dplyr::summarize(number_of_rides = n()) %>% 
  arrange(member_casual) %>% 
  ggplot(aes(x = member_casual, y = number_of_rides, fill = rideable_type)) +
  geom_col(position = "dodge")
```


Count the exact number of average_durations by ridable type
```{r echo=FALSE}
all_trips_data_v2 %>% 
  group_by(member_casual, rideable_type) %>% 
  dplyr::summarize(count=n())
```

As we can see from data above. Classic bike is the most popular bicycle type among the three bicycle type, followed by electric bike then docked bike.




#### average_durations of rides by ridable type

Run the data visualization of average_durations  by rideable_type and member type
```{r echo=FALSE}
all_trips_data_v2 %>% 
  group_by(member_casual, rideable_type) %>% 
  dplyr::summarize(average_duration = mean(ride_length_min)) %>% 
  arrange(member_casual) %>% 
  ggplot(aes(x = member_casual, y = average_duration, fill = rideable_type)) +
  geom_col(position = "dodge")

```


Count the exact number of average_durations by rideable_type and member type
```{r echo=FALSE}
all_trips_data_v2 %>% 
  group_by(member_casual, rideable_type) %>% 
  dplyr::summarize(ride_length_min = mean(ride_length_min))
```
From the data above, we can see that docked bike from casual riders are the highest minute ride. 
After we breakdown member type by each ride type, we can see for each ride type are casual riders are higher than annual member. 











### Rides by day of week
#### Number of rides by day of week and member type


Run the data visualization of number_of_rides by day_of_Week - member type
```{r echo=FALSE}
all_trips_data_v2 %>% 
  mutate(weekday = wday(started_at, label = TRUE)) %>% 
  group_by(member_casual, weekday) %>% 
  dplyr::summarize(number_of_rides = n()
                   ,average_duration = mean(ride_length_min)) %>%  
  arrange(member_casual, weekday) %>% 
  ggplot(aes(x = weekday, y = number_of_rides, fill = member_casual)) +
  geom_col(position = "dodge") 
```


Count the exact number of number_of_rides by day_of_Week - member type
```{r echo=FALSE}
all_trips_data_v2 %>% 
  group_by(member_casual, day_of_Week) %>% 
  dplyr::summarize(count=n())
```
We can see that the number of casual riders is lesser than the number of annual members on the weekday, and otherwise the weekends. 
From three data above, it is indicate there are different behavior between annual members and casual riders. 




#### average_durations of rides by day_of_Week- member_casual

Run the data visualization of average_durations by day_of_Week - member type
```{r echo=FALSE}
all_trips_data_v2 %>% 
  mutate(weekday = wday(started_at, label = TRUE)) %>% 
  group_by(member_casual, weekday) %>% 
  dplyr::summarize(number_of_rides = n()
                   ,average_duration = mean(ride_length_min)) %>%  
  arrange(member_casual, weekday) %>% 
  ggplot(aes(x = weekday, y = average_duration, fill = member_casual)) +
  geom_col(position = "dodge") 
```

Count the exact number of average_durations by day_of_Week - member type
```{r echo=FALSE}
all_trips_data_v2 %>% 
  group_by(member_casual, day_of_Week) %>% 
  dplyr::summarize(ride_length_min = mean(ride_length_min))
```

Difference behavior can also be seen from the ride length per minute, that casual riders have an average rides of around 16 to 29 minutes on weekdays, and higher on weekend of around 32 minutes to 35 minutes. 
Otherwise the annual members, Annual members have a driving record of 12 to 13 minutes on weekdays, and 14 to 15 minutes. Not too different every day for annual members.

I assume that even on weekdays casual riders who commute with bikes have a longer distance than annual members. But this assume need to be confirm from the distance on miles or km, and we didn't have the data to explore it.


Note : We can analyze more deeply with the average per user, using the bikes per day. "what is the total minutes per user per day. 











### Rides by trip_hour - member type
#### Number of rides by trip_hour - member type

Run the data visualization of number_of_rides by trip_hour - member type
```{r echo=FALSE}
all_trips_data_v2 %>% 
  group_by(member_casual, trip_hour) %>% 
  dplyr::summarize(number_of_rides = n()
                   ,average_duration = mean(ride_length_min)) %>%  
  arrange(member_casual, trip_hour) %>% 
  ggplot(aes(x = trip_hour, y = number_of_rides, fill = member_casual)) +
  geom_col(position = "dodge") +
  theme(axis.text.x = element_text(angle=90))
```

Count the exact number of number_of_rides by trip_hour - member type
```{r echo=FALSE}
all_trips_data_v2 %>% 
  group_by(member_casual, trip_hour) %>% 
  dplyr::summarize(count=n())

```
From data above, we can see that both of casual riders and annual members most of rides bicycle during evening time from 4pm - 6pm, but number gets lesser and lesser when the time close to 4pm in the morning. 
When morning peak hour 6 - 8 am, there is an increasing user for annual member. 





#### average_durations of rides by trip_hour - member type

Run the data visualization of average_durations by trip_hour - member type
```{r echo=FALSE}
all_trips_data_v2 %>% 
  group_by(member_casual, trip_hour) %>% 
  dplyr::summarize(number_of_rides = n()
                   ,average_duration = mean(ride_length_min)) %>%  
  arrange(member_casual, trip_hour) %>% 
  ggplot(aes(x = trip_hour, y = average_duration, fill = member_casual)) +
  geom_col(position = "dodge") +
  theme(axis.text.x = element_text(angle=90))
```


Count the exact number of average_durations by trip_hour - member type
```{r echo=FALSE}
all_trips_data_v2 %>% 
  group_by(member_casual, trip_hour) %>% 
  dplyr::summarize(ride_length_min = mean(ride_length_min))
```

Almost same result from the chart above, that average duration from casual riders are higher than annual member. 
Annual members have almost get same average duration around 11 - 14 minutes per one trip and per each hour. Otherwise, casual riders have a duration that varies from 19 to 37 minutes in one day.

This behavior difference can indicate that casual riders have a longer duration, due to longer rides compared to annual members.










### Rides by trip_hour - member type per each day_of_week (facet)
#### Number of rides by trip_hour - member type per each day_of_week (facet)

Run the data visualization of number_of_rides by day_of_Week - member type
```{r echo=FALSE}
all_trips_data_v2 %>% 
  group_by(member_casual, day_of_Week, trip_hour) %>% 
  dplyr::summarize(number_of_rides = n()
                   ,average_duration = mean(ride_length_min)) %>%  
  arrange(member_casual, trip_hour) %>% 
  ggplot(aes(x = trip_hour, y = number_of_rides, fill = member_casual)) +
  geom_col(position = "dodge") +
  facet_wrap(~day_of_Week) 
```

Count the exact number of average_durations
```{r echo=FALSE}
all_trips_data_v2 %>% 
  group_by(member_casual, day_of_Week, trip_hour) %>% 
  dplyr::summarize(count=n())
```

The trend is almost the same as the number of rides as what we saw previously for the weekday.The number of riders getting higher on morning peak hour and evening peak hours


Behavioral differences can be seen on weekends or weekdays, there are more annual members who rent the bicycle to commute to work on weekdays, and on weekend there are more casual riders who rent the bicycle to tour around, even though it looks like the number of them (casual riders and annual members) on weekends is not as much as weekdays











### Rides that has start and end bike in same station
#### Number of rides that has start and end bike in same station


Run the data visualization of number_of_rides that has start and end bike in same station
```{r echo=FALSE}
all_trips_data_v2 %>% 
  group_by(member_casual, start_end_same_station) %>% 
  dplyr::summarize(number_of_rides = n()
                   ,average_duration = mean(ride_length_min)) %>%  
  arrange(member_casual, start_end_same_station) %>% 
  ggplot(aes(x = start_end_same_station, y = number_of_rides, fill = member_casual)) +
  geom_col(position = "dodge")
```


Count the exact number of riders that has start and end bike in same station
```{r echo=FALSE}
all_trips_data_v2 %>% 
  group_by(member_casual, start_end_same_station) %>% 
  dplyr::summarize(count=n())

```

As we can see from bar chart above, that there are only a few riders who will return the bike on the same station where they rented the bike. It is mean that the riders usially will ride the bicycle to another place and return the bicycle there.






### Rides that has start and end bike in same station per day_of_week (facet)
#### Number of rides that has start and end bike in same station

Run the data visualization of number_of_rides that has start and end bike in same station per day_of_week (facet)
```{r echo=FALSE}
all_trips_data_v2 %>% 
  group_by(member_casual, day_of_Week, start_end_same_station) %>% 
  dplyr::summarize(number_of_rides = n()
                   ,average_duration = mean(ride_length_min)) %>%  
  arrange(member_casual, start_end_same_station) %>% 
  ggplot(aes(x = start_end_same_station, y = number_of_rides, fill = member_casual)) +
  geom_col(position = "dodge") +
  facet_grid(cols = vars(day_of_Week))
```


Count the exact number of rides that has start and end bike in same station per day_of_week (facet)
```{r echo=FALSE}
all_trips_data_v2 %>% 
  group_by(member_casual, start_end_same_station) %>% 
  dplyr::summarize(count=n())

```

We can see that there are more annual members who rent and bike the bicyle to another place and return the bicycle at another station on weekdays as compared the casual riders. 
The annual members who need to commute to work on weekdays will usually rent bicycle to cycle to the workplace and return bicycle at the station nearby the workplace.

While for weekends, the number of casual riders and annual members who rent bicycle are almost the same for returning bycyle at different station. 











### Rides by month by month and member type
#### Number of rides by trip_hour - member type

Run the data visualization of number_of_rides by month and member type
```{r echo=FALSE}
all_trips_data_v2 %>% 
  group_by(member_casual, month) %>% 
  dplyr::summarize(number_of_rides = n()
                   ,average_duration = mean(ride_length_min)) %>%  
  arrange(member_casual, month) %>% 
  ggplot(aes(x = month, y = number_of_rides, fill = member_casual)) +
  geom_col(position = "dodge") 
```


Count the exact number of rides by month and member type
```{r echo=FALSE}
all_trips_data_v2 %>% 
  group_by(member_casual, month) %>% 
  dplyr::summarize(count=n())
```

From bar chart above, shown us the total number of riders for each month. The number of riders start to increase after winter, or increase from February and reach the highest number on June, and then decrease again when hit the winter. 

Note : 
Winter : December to February
Spring : March to May
Summer : June to August
Autumn : September to November

#### total duration of rides by month and member type
```{r echo=FALSE}
all_trips_data_v2 %>% 
  group_by(member_casual, month) %>% 
  dplyr::summarize(number_of_rides = n()
                   ,average_duration = mean(ride_length_min)
                   ,total_duration = sum(ride_length_min)) %>%  
  arrange(member_casual, month) %>% 
  ggplot(aes(x = month, y = total_duration, fill = member_casual)) +
  geom_col(position = "dodge") 

```

Count the exact number of average_durations by month and member type
```{r echo=FALSE}
all_trips_data_v2 %>% 
  group_by(member_casual, month) %>% 
  dplyr::summarize(total_duration = sum(ride_length_min))
```

In bar chart above, we can see the total duration of riding for each month.
We can see that as winter approaches, there is a decrease in riders and after winter the riders start to increase.

This shows the behavior of riders when faced with several seasons, especially in winter, people may not use bicycles and will use other transportation.

And still consistent on the previous chart, that casual riders still have a higher total duration compared to annual members






Note: for recommendations to Cyclistic company and solutions to existing problems will be explained in power point